\chapter{The Case\label{case}}
% Johdanto
\section{Tiivistelmä}
This research studies performance of a server side JavaScript software component.
Component is written using JavaScript with Node.js.
The component performance is measured in two different environments.
We define performance as components response time.
Two environments are one where the component is a part of a monolith application.
Other where the components is running as an independent service.
The component is not modified between environments.

\section{Overview of the component}
The component is written in JavaScript using Node.js.
It's response time is critical to the application.
The component can not be scaled horizontally.

The component contains four modules.
Module A, B, C and D.
Some modules are dependent on external data.
Data fetching for these modules are excluded from the performance review.
Modules run in cycles.
One full cycle contains all modules from A to D.
Modules are run in sequence.

The \textbf{module A} is the first module in the sequence.
It starts the cycle.
It is always called.

The \textbf{module B} is the second module in the sequence.
It runs after \textbf{module A}.
It is always called.

The \textbf{module C} is the third module in the sequence.
It is run after \textbf{module B} only when response from \textbf{module B} triggers it.

The \textbf{module D} is the final module in the sequence.
It is run always after \textbf{module C} runs.

When any of the modules fails to perform its task an error is raised and the cycle is interrupted.
When modules are running without errors the modules A and B are always performed. 
The modules C and D are run only when the data from \textbf{module B} triggers them.
Each module waits for the response of every external call before continuing their process.

\section{Monolith overview}
The monolith uses Node.js version 14.19.1. The node version runs using V8 engine version \textit{8.4.371.23-node.85}.

The monolith application is a three tier application containing user interface, server side and database.
The server side contains four components.
One of the components is the component studied in this research.
The component is run independently from other components.

\section{The independent service}
The independent service uses Node.js version 14.18.3.
The node version runs using the v8 engine version \textit{8.4.371.23-node.85}.

The service contains the component studied in this research and some functionality that allows communication with the component.
